---
title: "Projections, joins, and extractions"
author: "Matt Williamson"
date: "1/27/2020"
output: 
    ioslides_presentation:
        css: style.css
        theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fasterize)
library(raster)
library(velox)
library(sf)
library(sp)
library(velox)
library(tigris)
library(tidyverse)
```

## Objectives
By the end of today, you should be able to:

- Access, understand, and change the projection of a `Spatial*` and `sf` object

- Join datasets by attribute or location

- Implement a workflow for extracting raster data to polygons (i.e., zonal stats)

## Projections
```{r projfig, echo=FALSE, fig.cap="Image Source: M. Corey, opennews.org", fig.height=1.5, fig.align="center"}
proj_path <- "./images/projections.png"
img1 <- png::readPNG(proj_path, native = TRUE, info = TRUE)
knitr::include_graphics(proj_path)
```
## Projections

- Projections tell your computer how to 'flatten' your data into a 2-dimensional map

- Goal is to choose projection that best maintains *both* distance and direction

- Coordinate Reference System (CRS) provides standardized way of describing location and projection
  - `R` stores this information as either a `proj4` string, an `EPSG` code
  - ESRI, ENVI, and a few other platforms also use `WKT`, the "Well-known text" format.
  - A CRS generally contains 


## Additional references
* Projections and Coordinate Reference Systems
  - [CRS library](https://spatialreference.org/ref/epsg/): searchable website for all of the EPSG and proj4 values for different projections
  
  - [Choosing the right map projection](https://source.opennews.org/articles/choosing-right-map-projection/): A deeper exploration of why projections matter and rules of thumb for selecting them.
  
  - [NCEAS cheatsheet on CRS data in R](https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/OverviewCoordinateReferenceSystems.pdf): A description of how projections are developed, some best practices, and examples using `sp` and `rgdal`.
