---
title: "Raster Operations"
author: "Matt Williamson"
date: "2/17/2020"
output: 
    ioslides_presentation:
        css: style.css
        theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(raster)
library(FedData)
library(tidyverse)
library(sf)
```

## Learning Objectives

- Explore basic raster operations (raster math, moving window analysis)

- Learn workflow for raster reclassification

- Learn basic approach for merging neighboring rasters

## A bit of remote sensing

- Deriving vegetation indices largely a function of calculating the right ratios.

- `R` treats `rasters` as large matrices making math pretty easy

```{r ldata, echo = TRUE}
if(!file.exists('data/rs/samples.rds')) {
  download.file('https://biogeo.ucdavis.edu/data/rspatial/rsdata.zip', 
                dest = 'data/rsdata.zip')
  unzip('data/rsdata.zip', exdir='data')}
```

## Load an image
```{r getdata, echo=TRUE}
raslist <-paste0('data/rs/LC08_044034_20170614_B', 1:11, ".tif")
landsat <- stack(raslist)
landsatRGB <- landsat[[c(4,3,2)]]
landsatFCC <- landsat[[c(5,4,3)]]
identical(crs(landsat), crs(landsatRGB))
identical(crs(landsatRGB), crs(landsatFCC))
```

## Plot the image
```{r lsplot, echo=TRUE, warning=FALSE}
par.orr <- par()
par(mfrow =c(1,2))
plotRGB(landsatRGB, axes=TRUE, stretch="lin", main="Landsat True Color Composite")
plotRGB(landsatFCC, axes=TRUE, stretch="lin", main="Landsat False Color Composite")
par(par.orr)
```

## Calculate Normalized Difference Vegetation Index

- NDVI is a commonly used expression of "greenness" used to characterize productivity

- Operationally just a series of cell-wise arithmetic
```{r ndvimath, echo=TRUE}
#Landsat NIR = band 5, red = 4
NDVI.numerator <- landsat[[5]] - landsat[[4]] 
NDVI.denominator <- landsat[[5]] + landsat[[4]]
NDVI <- NDVI.numerator/NDVI.denominator
```

## Plot NDVI
```{r}
plot(NDVI, col=rev(terrain.colors(20)), main = "NDVI")
```

## Raster math continued

- More complex operations possible

- rescaling data

```{r centerndvi, echo=TRUE}
NDVI.centered <- NDVI - cellStats(NDVI, "mean")
plot(NDVI.centered, col=rev(terrain.colors(20)), main = "NDVI (centered)")
```

## Manual Scaling
```{r scalendvi2, echo=TRUE}
NDVI.scaled <- (NDVI - cellStats(NDVI, "mean"))/cellStats(NDVI,"sd")
plot(NDVI.scaled, col=rev(terrain.colors(20)), main = "NDVI (scaled)")
```

## Using `scale` in `raster`

- `raster` package recreates a number of generic functions for `Raster` objects
```{r scalendvi, echo=TRUE}
NDVI.scaled2 <- scale(NDVI)
plot(NDVI.scaled2, col=rev(terrain.colors(20)), main = "NDVI (scaled)")
```

## Reclassifying rasters

- Useful for converting continuous rasters into categorical (or ordinal) data

- Thresholding based on a minimum value

```{r vegclass, echo=TRUE}
veg <- reclassify(NDVI,cbind(-Inf, 0.4,NA))
plot(veg, col=rev(terrain.colors(20)), main = "Vegetation")
```

## Reclassifying (cont'd)

- Based on percentiles in the data
```{r percreclass, echo=TRUE}
ndvi.percentile <- quantile(NDVI, probs = seq(0,1, 0.1), names=FALSE)

#build a reclassification matrix
perc.rcl.df <- NULL
for (i in 1:length(ndvi.percentile)){
  perc.rcl.v <- c(ndvi.percentile[i], ndvi.percentile[i+1], i)
  perc.rcl.df <- rbind(perc.rcl.df, perc.rcl.v)
  rownames(perc.rcl.df) <- NULL
}
perc.rcl.df[11,2] <- Inf #set the upper limit

ndvi.perc.recl <- reclassify(NDVI, perc.rcl.df) 
```

## Reclassified

```{r reclplot, echo=TRUE}
plot(ndvi.perc.recl, col=rev(terrain.colors(20)), main = "Percentiles")
```


## Reclassifying categorical maps

- Useful for land cover, vegetation, etc

- Need to create a lookup table for `reclassify`

```{r echo=FALSE, message=FALSE}
#download NLCD data from MRLC website
nlcdAB <- raster("data/nlcd_adaboise.tif")

AB <- c("Ada", "Boise")

AB.poly <- tigris::counties(st = "ID") %>%
  as(.,"sf") %>% 
  filter(NAME %in% AB) %>% 
  st_transform(., crs = proj4string(nlcdAB)) %>% 
  as(., "Spatial")
nlcdAB <- crop(nlcdAB, AB.poly)
```

## NLCD
<<<<<<< HEAD
=======

>>>>>>> 42ac145af02b1ca7d240b4ca86db65e6866b2208
```{r nlcdplot, echo=FALSE}
plot(nlcdAB)
plot(AB.poly, border = "blue", add = TRUE)
```

## NLCD reclassified

- Let's reclassify the developed categories 

- Create a lookup table based on the original values in the NLCD

- use `grepl` to find and replace the developed categories

```{r nlcdclass, echo=TRUE}
original.val <- unique(nlcdAB)
#reclassify developed categories
reclass.val <- ifelse(grepl("^2", original.val), "21", original.val) 

reclass.tabl <- matrix(c(as.numeric(original.val), as.numeric(reclass.val)), ncol = 2)

nlcd.reclass <- reclassify(nlcdAB, reclass.tabl)
```

## Reclassified NLCD

```{r pltrecls}
plot(nlcd.reclass,main="NLCD reclassified")
```


## Mosaicing rasters

- Lots of satellite imagery comes in "tiles"

- Speeding up processing often requires "tiling"

- Mosaicing allows us to put them back together


## Using merge to combine NLCD 

- when objects have the same origin and extent you can use `raster::merge`

- in areas where objects overlap, priority is based on the order of the arguments

- setting the `tolerance` option can relax the origin and extent enforcement

```{r plotnlcd}

nlcdCGO <- raster("data/nlcd_canyongemowyhee.tif")
tv.1 <- c("Canyon", "Gem", "Owyhee")
tv.cty.1 <- tigris::counties(st = "ID") %>%
  as(.,"sf") %>% 
  filter(NAME %in% tv.1) %>% 
  st_transform(., crs = proj4string(nlcdCGO)) %>% 
  as(., "Spatial") 
nlcdCGO <- crop(nlcdCGO,tv.cty.1)

```

## The `merged` NLCD dataset for the Treasure Valley

```{r nlcdmerge, echo=TRUE}
nlcdmerge <- merge(nlcdAB, nlcdCGO)
plot(nlcdmerge)
```

## Using `raster::mosaic`

- similar to `merge`

- Allows you to use a function (e.g., mean, max, min) when cells overlap

- can use `do.call` when you have a large list of rasters

```{r mos, echo=TRUE}
nlcd.mosaic <- mosaic(nlcdAB, nlcdCGO, fun = min)

nlcd.list <- list(nlcdAB, nlcdCGO)
names(nlcd.list)[1:2] <- c('x', 'y')
nlcd.list$fun <- mean
nlcd.list$na.rm <- TRUE

nlcd.mos <- do.call(mosaic, nlcd.list)
```

## NLCD mosaic

```{r plotmos}
tv.1 <- c("Canyon", "Gem", "Owyhee")
tv.2 <- c("Ada", "Boise")

tv.cty.1 <- tigris::counties(st = "ID") %>%
  as(.,"sf") %>% 
  filter(NAME %in% tv.1) %>% 
  st_transform(., crs = proj4string(nlcdCGO)) %>% 
  as(., "Spatial") 

tv.cty.2 <- tigris::counties(st = "ID") %>%
  as(.,"sf") %>% 
  filter(NAME %in% tv.2) %>% 
  st_transform(., crs = proj4string(nlcdAB)) %>% 
  as(., "Spatial") 

plot(nlcd.mos)
plot(tv.cty.1, border = "blue", add = TRUE)
plot(tv.cty.2, border = "red", add = TRUE)

```

## Addtional resources

[Robert Hijmans and Ani Ghosh Remote Sensing Book](https://rspatial.org/rs/rs.pdf)

[The landsat package](https://cran.r-project.org/web/packages/landsat/landsat.pdf)

[The getlandsat package](https://docs.ropensci.org/getlandsat/)