---
title: "Projections, joins, and extractions"
author: "Matt Williamson"
date: "1/27/2020"
output: 
    ioslides_presentation:
        css: style.css
        theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fasterize)
library(raster)
library(velox)
library(sf)
library(sp)
library(velox)
library(tigris)
library(tidyverse)
library(maps)
library(mapproj)
library(patchwork)
```

## Objectives
By the end of today, you should be able to:

- Access, understand, and change the projection of a `Spatial*` and `sf` object

- Join datasets by attribute or location

- Implement a workflow for extracting raster data to polygons (i.e., zonal stats)

## Projections
```{r projfig, echo=FALSE, fig.cap="Image Source: M. Corey, opennews.org", fig.height=1.5, fig.align="center"}
proj_path <- "./images/projections.png"
img1 <- png::readPNG(proj_path, native = TRUE, info = TRUE)
knitr::include_graphics(proj_path)
```

## Projections
```{r usproj, echo=FALSE, fig.align="top", fig.height=3}
us_states <- map_data("state")
usamap1<-ggplot(us_states, aes(x=long, y=lat, group=group)) +
  geom_polygon(fill="white", colour="black") + 
  coord_map("mercator") +
  ggtitle("Mercator")
usamap2 <- ggplot(us_states, aes(x=long, y=lat, group=group)) +
  geom_polygon(fill="white", colour="black") + 
  coord_map("azequalarea") +
  ggtitle("Equal Area")
usamap1 + usamap2
```


- Projections tell your computer how to 'flatten' your data into a 2-dimensional map

- Goal is to choose projection that best maintains *both* distance and direction

## The Coordinate Reference System

- Coordinate Reference System (CRS) provides standardized way of describing location and projection
  - `R` stores this information as either a `proj4` string, an `EPSG` code
  
  - ESRI, ENVI, and a few other platforms also use `WKT`, the "Well-known text" format.

## Anatomy of a Coordinate Reference System

- A CRS generally contains:
  
  - the Ellipsoid: Describes the general shape of the globe (not exactly spherical)
  
  - the Datum: Defines the origin and orientation of the coordinate access
  
  - the Projection: Projects the globe onto a 2-D surface
  
  - the units: the units of the map (e.g., meters, decimal degrees, arc-seconds)
  
  - the conversion used to convert to the ellipsoid

## Accessing the CRS in R
```{r crsdiscplay, eval=FALSE}
idaho <- rgdal::readOGR("./data/idaho_cty.shp")
sp::proj4string(idaho)

idaho.sf <- sf::st_read("./data/idaho_cty.shp")
sf::st_crs(idaho.sf)

```

## Re-projections in `sp`
```{r reprojsp, eval=FALSE}
#choose a few projections
nad83 <- "+init=epsg:4269"
idaho.cent <- "+proj=tmerc +lat_0=41.66666666666666 +lon_0=-114 +k=0.9999473679999999 +x_0=500000.0001016001 +y_0=0 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs " 

idaho.83 <- spTransform(idaho, CRS(nad83))
idaho.c <- spTransform(idaho, CRS(idaho.cent))

#are they different?
identicalCRS(idaho.83, idaho.c)

#reproject idaho.cent
idaho.reproj <- spTransform(idaho.83, proj4string(idaho.c))

#are they different?
identicalCRS(idaho.c, idaho.reproj)
```
## Re-projections in `sf`
```{r reprojsf, eval=FALSE}
#choose a few projections
nad83 <- "+init=epsg:4269"
idaho.cent <- "+proj=tmerc +lat_0=41.66666666666666 +lon_0=-114 +k=0.9999473679999999 +x_0=500000.0001016001 +y_0=0 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs " 

idaho.sf.83 <- sf::st_transform(idaho.sf, nad83)
idaho.sf.c <- sf::st_transform(idaho.sf, idaho.cent)
## are they the same?
st_crs(idaho.sf.83) == st_crs(idaho.sf.c)

##reproject using sf syntax
idaho.sf.reproj <- st_transform(idaho.sf.83, st_crs(idaho.sf.c))

## are they the same?
st_crs(idaho.sf.reproj) == st_crs(idaho.sf.c)

```


## Joining attributes to spatial data

- Sometimes we want to add addtional attributes to our vector data

- We can "Join" data together by shared attributes or locations

  - inner join: all rows from df1 with matching values in df2
  
  - left join: all rows from df1 and all columns from df1 and df2 (NAs where no match)
  
  - right join: all rows from df2 and all columns from df1 and df2

## Joins continued  
  - full join: all rows and columns from both dfs
  
  - semi join: all rows from df1 with matching values in df2, only columns from df1
  
  - anti join: all rows from df1 without matching values in df2, only columns from df2

## Joining by attributes using `sp` and `tigris`
```{r spjoin, eval=FALSE}
head(idaho.83@data)

#load a tabular dataset
idaho.nonprof <- read.csv("./data/bmfcount.csv", stringsAsFactors = FALSE, colClasses = "character")
head(idaho.nonprof)

#join based on the columns that match
idaho.join <- tigris::geo_join(idaho.83, data_frame = idaho.nonprof, by_sp = "GEOID", by_df="FIPS", how="left")
head(idaho.join@data)
str(idaho.join@data)
```


## Joining by attrbutes using `sf`
```{r}
head(idaho.sf.83)
str(idaho.sf.83)

idaho.join.sf <- idaho.sf.83 %>% 
  left_join(., idaho.nonprof, by = c("GEOID" = "FIPS"))
head(idaho.join.sf)
str(idaho.join.sf)


```

## Join by location in 'sp'

- can use `over` from `sp`

- Sometimes strange behavior when more than one intersection

- Can also use `gIntersects` from `rgeos` or `union` from `raster` but syntax bcomes more complicated

```{r spatialjoinsp, eval=FALSE}
#get some places within counties
idaho.places <- tigris::places(state="16", year = "2015")

#Ar the projections the same
identicalCRS(idaho.join, idaho.places)

#need to reproject idaho.places
idaho.places.proj <- spTransform(idaho.places, proj4string(idaho.join))

#join based on location
nonprof.join <- cbind(over(idaho.join, idaho.places.proj), as.data.frame(idaho.join))

str(nonprof.join)

```


## Additional references
* Projections and Coordinate Reference Systems
  - [CRS library](https://spatialreference.org/ref/epsg/): searchable website for all of the EPSG and proj4 values for different projections
  
  - [Choosing the right map projection](https://source.opennews.org/articles/choosing-right-map-projection/): A deeper exploration of why projections matter and rules of thumb for selecting them.
  
  - [NCEAS cheatsheet on CRS data in R](https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/OverviewCoordinateReferenceSystems.pdf): A description of how projections are developed, some best practices, and examples using `sp` and `rgdal`.
