---
title: "Introduction to spatial datatypes"
author: "Matt Williamson"
date: "1/20/2020"
output: 
  ioslides_presentation:
        css: style.css
        theme: readable
---
<style>
p.caption {
  font-size: 0.6em;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

## Objectives
By the end of today, you should be able to:

- Distinguish amongst 3 common spatial data-types

- Identify the "major" packages for dealing with each datatype

- Recognize the differences between the objects that each package creates

# Vector Data

## Vector Data


## Vector Data

```{r vecfig, echo=FALSE, fig.cap="Image Source: Colin Williams (NEON)", fig.height=1.5}
img1_path <- "./images/points-lines-polygons-vector-data-types.png"
img1 <- png::readPNG(img1_path, native = TRUE, info = TRUE)
knitr::include_graphics(img1_path)
```

## Vector Data

* **Points:** Points defined by a single x, y coordinate. There can be many points in a vector point file. 
    + *Examples:* sampling locations, control points, addresses

* **Lines:** Lines are composed of many (at least 2) vertices, or points, that are connected. The more vertices, the more complex (i.e., larger in terms of memory) the line.
    + *Examples:* Roads, rivers, transects


* **Polygons:** A polygon consists of 3 or more vertices that are connected and “closed”. Note that self-intersections and holes can lead to invalid geometries that disrupt spatial analyses.
    + *Examples:* state boundaries, sample sites, buffered transects 


## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

