---
title: "Welcome to Space!"
subtitle: "HES 598 - Introduction to Spatial Data in R"
author: "Matt Williamson"
date: "1/14/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introductions

- A bit about me
- A bit about you
- [Course Webpage](https://hes-bsu.github.io/HES598/)
- Git, Github, etc...

## The Syllabus

- Expectations (Yours and Mine)
- Course Design
- Advanced Topics

## Why R?

- Open Source
- Integrated analysis pipelines
- Reproducible workflows 

## Why Not R?

- Coding can be hard...
- Memory challenges
- Speed
- Decision fatigue

## What is spatial data?

## 
```{r echo=FALSE}
library(maps)
library(socviz)
library(tidyverse)
party_colors <- c("#2E74C0", "#CB454A") 

us_states <- map_data("state")
head(us_states)
election$region <- tolower(election$state)
us_states_elec <- left_join(us_states, election)
p0 <- ggplot(data = us_states_elec,
             mapping = aes(x = long, y = lat,
                           group = group, fill = party))
p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 
p2 <- p1 + scale_fill_manual(values = party_colors) +
    labs(title = "Election Results 2016", fill = NULL)
p2
```

##
```{r}
p0 <- ggplot(data = subset(election, st %nin% "DC"),
             mapping = aes(x = r_points,
                           y = reorder(state, r_points),
                           color = party))

p1 <- p0 + geom_vline(xintercept = 0, color = "gray30") +
    geom_point(size = 2)

p2 <- p1 + scale_color_manual(values = party_colors)

p3 <- p2 + scale_x_continuous(breaks = c(-30, -20, -10, 0, 10, 20, 30, 40),
                              labels = c("30\n (Clinton)", "20", "10", "0",
                                         "10", "20", "30", "40\n(Trump)"))

p3 + facet_wrap(~ census, ncol=1, scales="free_y") +
    guides(color=FALSE) + labs(x = "Point Margin", y = "") +
    theme(axis.text=element_text(size=8))
```

## Things to think about
- Am I making a map or running an analysis
- What is the goal of my analysis or figure?
- Is the spatial arrangement important?