---
title: "Introduction to Visualization"
author: "Matt Williamson"
date: "2/1/2020"
output: 
    ioslides_presentation:
        css: style.css
        theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=60), tidy = TRUE, message = FALSE)
library(tidycensus)
library(tidyverse)
library(sf)
library(raster)
library(tmap)
library(rasterVis)
library(latticeExtra)
library(RColorBrewer)
library(viridis)
library(tigris)
```

## Today's objectives
- Explore plotting vector data using `plot`, `spplot`, and `ggplot2`

- Explore plotting raster data using `plot`, `rasterViz`

- Map multiple layers using `spplot`, `ggplot2`


## Building a vector dataset

```{r}
census_api_key("90b94953d2f24e81e890229e0128174f5ba80d3f")
#download censu data
id.income <- get_acs(geography = "county", 
              variables = c("B19013_001", "B01002_001"),
              state = "ID", 
              year = 2018, geometry = TRUE) %>% 
  mutate(rename = 
           str_replace_all(.$variable,
                           (c("B19013_001"="income",
                                         "B01002_001"="age" )))) %>%
  #select the columns to keep
  dplyr::select(c(GEOID, rename, estimate)) %>% 
  #have to use spread because pivot_wider 
  #not currently working for sf objects
  spread(., key = rename, value = estimate) 

#join our nonprofit data set
id.nonprof <- read.csv("./data/bmfcount.csv", 
                       stringsAsFactors = FALSE, 
                       colClasses = "character")

id.census.sf <- id.income %>% 
  left_join(., id.nonprof, by = c("GEOID" = "FIPS"))

id.census.sp <- as(id.census.sf, "Spatial")
```


## Plotting spatial data in base `plot`

- Base `plot` handles `Spatial*` and `sf` objects differently
- Base `plot` is useful for quick exploration of data and figure drafting

## Plotting spatial data in base `plot`

```{r vectorplot, echo = TRUE}
plot(id.census.sp, main = "Idaho Counties", 
     sub="2018 Census Data", 
     col="blue", 
     border="white")


```

## Plotting `sf` data in base `plot`

```{r}
plot(id.census.sf, main= "Idaho Counties", 
     sub="2018 Census Data", 
     col="blue", 
     border="white")
```

## Plotting `sf` data in base `plot`

```{r}
#if you just want the sf geometry
plot(st_geometry(id.census.sf), 
     col="blue", 
     border="white")
```

## Plotting attributes of `Spatial*` data in base `plot`
```{r}
#hard to plot by attributes in base plot with Spatial* objects
plot(id.census.sp$income)
#Easier with sf

```

## Plotting attributes of `sf` data in base `plot`

```{r}
plot(id.census.sf["income"], 
     main= "Idaho Income", 
     sub="Median county value as of 2018")
```

## Plotting raster data in baseplot

```{r rasterplot}
id.elev <- raster("data/id_elev.tif")

```

## Plotting raster `bricks` in baseplot

```{r}
#create a reaster brick
id.terrain <- terrain(id.elev, 
                      opt=c('slope', 'aspect','TRI', 'TPI'))
plot(id.terrain)
```

## Adding functionality with `spplot`
- `spplot` follows the `lattice` approach for creating graphics

- Allows functionality for `Spatial*` objects similar to those for `sf` in base `plot`

```{r}
#set up a color palette
income.pal <- brewer.pal(n = 7, name = "Greens")
spplot(id.census.sp, 
       col.regions = income.pal, 
       cuts = 6)

```

## Plotting a single variable in `spplot`

```{r}
spplot(id.census.sp, 'income', 
       col.regions = income.pal, 
       cuts = 6)
```

## Adding layers 

- Rare that we only want to visualize one aspect of spatial data

- Adding layers allows additional information and detail

- Important to ask "is a map the best way to display this?"

## Adding layers in `spplot`

- add layers using `list` and `sp.layout`

- order matters!!

## Adding layers in `spplot`

```{r, spplotlyr}
county.layer <- list("sp.polygons", id.census.sp, 
                     fill="transparent", 
                     col = "blue", 
                     lwd = 1.5, first=FALSE)

spplot(id.census.sp, 'income', 
       col.regions = income.pal, 
       cuts = 6, 
       sp.layout = county.layer)
```

## Adding rasters in `spplot`

```{r}
spplot(id.elev, sp.layout = county.layer)
```

## Adding rasters in `spplot`

```{r}
spplot(id.terrain, sp.layout = county.layer)
```

## Adding map elements with `spplot`

- You'll often want additional components for maps

- These can also be added with `sp.layout`

- Need to specify size and coordinates for elements

## Adding map elements with `spplot`

```{r scale}
id.proj <- spTransform(id.census.sp, CRS("+init=EPSG:32611"))  # Can't make a scale if not projected!
id.proj@bbox  # Check dimensions to help guide offset choices
#expand bbox to make room for map elements
new.bbox <- matrix(c(480000, 1010000, 4600000, 5450000), 
                   ncol=2, byrow = TRUE, 
                   dimnames = list(c("x", "y"),c( "min", "max")))
id.proj@bbox <- new.bbox
bbox(id.proj) #check new bbox

county.layer <- list("sp.polygons", id.proj, 
                     fill="transparent", 
                     col = "green", 
                     lwd = 1.5, first=FALSE)

scale <- list("SpatialPolygonsRescale", 
              layout.scale.bar(), scale = 100000, 
              fill = c("transparent", "black"), 
              offset = c(490500, 4615000), 
              first=FALSE)

# The scale argument sets length of bar in map units
text1 = list("sp.text", c(490500, 4604800), "0", 
             cex =0.5, first=FALSE)
text2 = list("sp.text", c(590500, 4604800), "100 km", 
             cex=0.5, first=FALSE)

arrow = list("SpatialPolygonsRescale", layout.north.arrow(), 
             offset = c(980000, 5400000), 
             scale = 60000, first=FALSE)
```

## Adding map elements with `spplot`

```{r}
spplot(id.proj, "income", 
       sp.layout = 
         list(county.layer,scale, text1, text2, arrow))
```

## Using `rasterViz` for plotting multiple rasters

- `rasterViz` also builds on `lattice` approach

- lots of functionality for space, time, and spacetime data

- especially useful for layering `rasters`

## Using `rasterViz` for plotting multiple rasters

```{r}
#build a hillshade
id.hills <- hillShade(id.terrain[[1]], 
                      id.terrain[[2]], 
                      angle=40, 
                      direction = 270)
hills <- levelplot(id.hills, par.settings = GrTheme())
#crop raster - make sure projections match!
elev.crop <- mask(id.elev, id.census.sp)
elev.lp <- levelplot(elev.crop)
#combine rasters
hills + elev.lp
```


## Fancier graphics with `ggplot2`

- `ggplot2` built on grammer of graphics

- Plots build in layers denoted by different `geom_`, `aes`, and `scale_`

- Lots of functionality, but slow

## Fancier graphics with `ggplot2`

```{r}

#get tract-level income data
tv.income.tct <- get_acs(geography = "tract", 
              variables = "B19013_001",
              state = "ID", 
              year = 2018, geometry = TRUE) %>% 
  filter(str_detect(NAME, "Ada | Boise | Canyon | Gem | Owyhee") )

cty.map <- ggplot() +
  geom_sf(data = id.census.sf, fill = "gray")


```

## Adding layers in `ggplot2`

```{r}
cty.map +
  geom_sf(data = tv.income.tct, aes(fill = estimate)) +
  scale_color_viridis()
```

## Is this a reasonable way to look at this data?

```{r}
med.inc <- median(id.census.sf$income)

ggplot() +
  geom_point(data=id.census.sf, 
             aes(x=income, 
                 y = reorder(GEOID, income))) +
  geom_vline(xintercept = med.inc)
```


## Looking Ahead
- Plotting in R can be slow, we'll explore options in "Repetitive tasks, functional programming, and bench-marking"

- We will explore base-maps, multi-panel graphics, combined plots, and more complex layouts in "Next Level Visualization"

- We can explore interactive and web-based maps using the `leafletR` and `tmap` packages in "Flex Dashboards, web mapping, and interactive maps"



## Additional references
[Data Visualization: A Practical Introduction by Kieran Healy](https://socviz.co)

[The rasterViz package](https://oscarperpinan.github.io/rastervis/#levelplot)

[Plotting Simple Features](https://r-spatial.github.io/sf/articles/sf5.html)
