---
pagetitle: "HES 598 Syllabus"
title: "HES 598: Intro to Spatial Data in R"
date: "Updated: `r format(Sys.Date())`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars, echo=FALSE}
mon <- as.Date("2020-01-14")

advdate <- function(obj, adv) {
 tmon <- obj + 7*(adv-1)
 tfri <- obj + 4 + 7*(adv-1)
 tmon <- format(tmon, format="%m/%d")
 tfri <- format(tfri, format="%m/%d")
 zadv <- sprintf("%02d", adv)
 tmp <- paste("Week ",zadv,sep='',", ", tmon)
 return(tmp)
}
```
## Instructor Details
Dr. Matt Williamson

Human Environment Systems

College of Innovation and Design

[Environmental Research Building 4123](https://goo.gl/maps/RNBM9Y92P6FSuVgM8)

Office Hours: By appointment ([email](mailto:mattwilliamson@boisestate.edu))

## Course Description

Spatial data are ubiquitous and form the basis for many of our inquiries into social, ecological, and evolutionary processes. As such, developing the skills necessary for incorporating spatial data into reproducible statistical workflows is critical. In this course, we will introduce the core components of manipulating spatial data within the R statistical environment including managing vector and raster data, projections, extraction of data values, interpolation, and plotting. Students will also learn to prototype and benchmark different workflows to aid in applying the appropriate tools to their research questions.

## Course Objectives

Students completing this course should be able to:

 - Introduce commonly encountered spatial datatypes (polygons, rasters, netCDF)
 - Explore the variety of packages available for conducting spatial analyses in R
 - Introduce functional programming and benchmarking as a means of evaluating different approaches for managing spatial data
 - Develop workflows for extracting spatial data for use in statistical analyses
 - Practice reproducible workflow and visualization using Rmarkdown


## Basic Course Structure
This course will consist primarily of short (20-25 min) lectures followed by some in-class coding to help guide your exploration of the tools. I will lead the first half of the course to help get us all to a reasonable starting place. After that, the course is _yours_ to design and deliver. The goal here is to learn the spatial skills that are useful to _your_ research. I'll be here to help support and guide you, but you will get out of the course what you put into it once we move into the *Advanced Topics* portion of the course.

## Advanced Topics
This course is designed to help support your spatial analysis needs. As such, you are integral part to designing and delivering the parts of the course that are most useful to you. During the *Advanced Topics* portion of the course, I will work with you and your partner to develop a presentation that you will deliver to your classmates. We will try to describe the problem you are trying to solve. We will work together to identify at least two potential solutions that you will evaluate using the tools we learn in class. We will then convert each presentation into a blog post that can be kept on the course website as a resource to you and future students.

## Expectations
Seminar courses like these are only successful if we all work together. Shared expectations are a big part of that. As such, I expect the following from you:

1. Participation - I expect you to show up, ask questions, work through the in-class coding exercises, and be generally present while we are together.

2. Understanding - We are all at different places in our research and our experience with scientific computing. Rather than belittle each other for not knowing how to do something, I hope to create an environment where we all learn from each other and there are no stupid questions. I expect you to help me achieve this.

3. Communication - I am happy to shift directions in the course if it stops meeting your needs. I can only do that if you let me know that a change is necessary. Please feel free to provide feedback early and often.

You can expect the following from me:

1. Availability - This is a 1 credit course and your questions are likely to require more time than we can devote in class. I will do my best to make myself available outside of class hours to help you work through the spatial analysis problems that are slowing down your research progress.

2. Balance - The goal of this course is to teach you general approaches and workflows for spatial analyses in R, not to troubleshoot individual coding problems. I will do my best to point you toward resources that can help you solve your individual problems without monopolizing the time of others in class.

## Course Schedule
###  `r advdate(mon, 1)`: Introduction to spatial data and R

###  `r advdate(mon, 2)`: Vectors, rasters, and other types of spatial data

###  `r advdate(mon, 3)`: Finding alignment and extracting data

###  `r advdate(mon, 4)`: Introduction to visualization

###  `r advdate(mon, 5)`: No Class - complete R markdown exercise

###  `r advdate(mon, 6)`: Reclassify, mosaic, and other raster operations

###  `r advdate(mon, 7)`: Repetitive tasks, functions, and bench-marking

###  `r advdate(mon, 8)`: Intro to parallel computing for spatial data

###  `r advdate(mon, 9)`: Advanced topics: _Study design and spatial simulation_ 

###  `r advdate(mon, 10)`: Advanced topics: _Spatial stats, point patterns, nearest neighbors_

###  `r advdate(mon, 11)`: No Class (Spring Break)

###  `r advdate(mon, 12)`: Advanced topics: _Landscape metrics and raster data_

###  `r advdate(mon, 13)`: Advanced topics: _Telemetry and connectivity_

###  `r advdate(mon, 14)`: Advanced topics: _Species distribution models, statistical learning, interpolation_

###  `r advdate(mon, 15)`: Advanced topics: _Next-level visualization (publication quality figures, network maps, combining datatypes)_

###  `r advdate(mon, 16)`: Flex dashboards, web mapping, interactive maps

