---
pagetitle: "HES 598 Syllabus"
title: "HES 598: Intro to Spatial Data in R"
date: "Updated: `r format(Sys.Date())`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars, echo=FALSE}
mon <- as.Date("2020-01-14")

advdate <- function(obj, adv) {
 tmon <- obj + 7*(adv-1)
 tfri <- obj + 4 + 7*(adv-1)
 tmon <- format(tmon, format="%m/%d")
 tfri <- format(tfri, format="%m/%d")
 zadv <- sprintf("%02d", adv)
 tmp <- paste("Week ",zadv,sep='',", ", tmon)
 return(tmp)
}
```
## Course Description

Spatial data are ubiquitous and form the basis for many of our inquiries into social, ecological, and evolutionary processes. As such, developing the skills necessary for incorporating spatial data into reproducible statistical workflows is critical. In this course, we will introduce the core components of manipulating spatial data within the R statistical environment including managing vector and raster data, projections, extraction of data values, interpolation, and plotting. Students will also learn to prototype and benchmark different workflows to aid in applying the appropriate tools to their research questions.

 

## Course Objectives

Students completing this course should be able to:

 - Introduce commonly encountered spatial datatypes (polygons, rasters, netCDF)
 - Explore the variety of packages available for conducting spatial analyses in R
 - Introduce functional programming and benchmarking as a means of evaluating different approaches for managing spatial data
 - Develop workflows for extracting spatial data for use in statistical analyses
 - Practice reproducible workflow and visualization using Rmarkdown


## Expectations

## Advanced Topics
This course is designed to help support your spatial analysis needs. As such, you are integral part to designing and delivering the parts of the course that are most useful to you. During the *Advanced Topics* portion of the course, you will partner with one of your classmates and work with me to develop a presentation to your classmates. We will be convert each presentation into a blog post that can be kept on the course website as a resource to you and future students.

## Course Schedule
###  `r advdate(mon, 1)`: Introduction to spatial data and R

###  `r advdate(mon, 2)`: Vectors, rasters, and other types of spatial data

###  `r advdate(mon, 3)`: Finding alignment and extracting data

###  `r advdate(mon, 4)`: Introduction to visualization

###  `r advdate(mon, 5)`: No Class - complete R markdown exercise

###  `r advdate(mon, 6)`: Reclassify, mosaic, and other raster operations

###  `r advdate(mon, 7)`: Repetitive tasks, functional programming, and bench-marking

###  `r advdate(mon, 8)`: Intro to parallel computing for spatial data

###  `r advdate(mon, 9)`: Advanced topics: Study design and spatial simulation 

###  `r advdate(mon, 10)`: Advanced topics: Spatial stats, point patterns, nearest neighbors

###  `r advdate(mon, 11)`: No Class (Spring Break)

###  `r advdate(mon, 12)`: Advanced topics: Landscape metrics and raster data

###  `r advdate(mon, 13)`: Advanced topics: Telemetry and conectivity

###  `r advdate(mon, 14)`: Advanced topics: Species distribution models, geographically weighted regression, interpolation

###  `r advdate(mon, 15)`: Advanced topics: Next-level visualization (publication quality figures, network maps, combining datatypes)

###  `r advdate(mon, 16)`: Flex dashboards, web mapping, interactive maps

