---
pagetitle: "HES 598 Syllabus"
title: "HES 598: Introduction to Spatial Data in R"
date: "Updated: `r format(Sys.Date())`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars, echo=FALSE}
mon <- as.Date("2020-01-14")

advdate <- function(obj, adv) {
 tmon <- obj + 7*(adv-1)
 tfri <- obj + 4 + 7*(adv-1)
 tmon <- format(tmon, format="%m/%d")
 tfri <- format(tfri, format="%m/%d")
 zadv <- sprintf("%02d", adv)
 tmp <- paste("Week ",zadv,sep='',", ", tmon)
 return(tmp)
}
```

###  `r advdate(mon, 1)`: Introduction to spatial data and R

###  `r advdate(mon, 2)`: Vectors, rasters, and other types of spatial data

###  `r advdate(mon, 3)`: Finding alignment and extracting data

###  `r advdate(mon, 4)`: Introduction to visualization

###  `r advdate(mon, 5)`: No Class - complete R markdown exercise

###  `r advdate(mon, 6)`: Reclassify, mosaic, and other raster operations

###  `r advdate(mon, 7)`: Repetitive tasks, functional programming, and bench-marking

###  `r advdate(mon, 8)`: 

###  `r advdate(mon, 9)`: Advanced topics:

###  `r advdate(mon, 10)`: Advanced topics:

###  `r advdate(mon, 11)`: No Class (Spring Break)

###  `r advdate(mon, 12)`: Advanced topics:

###  `r advdate(mon, 13)`: Advanced topics:

###  `r advdate(mon, 14)`: Advanced topics:

###  `r advdate(mon, 15)`: Advanced topics:

###  `r advdate(mon, 16)`: Advanced topics:

