---
title: "Advanced Topic - Shiny Application Development"
author: "Lauren Hunt"
date: "3/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduce your problem

Give a brief introduction to your research and identify how your "Advanced Topic" plays into what you are hoping to do. For example, if you are just starting out on your project and your topic is "Study Design" you might be interested in developing a spatially informed sampling/stratification scheme.Or maybe your work is going to have a big public outreach component that you'd like to support with some nice interactive maps. Give us a sense for what "success" would look like - what is your ideal end product for the analysis/approach you're testing out. This will help us figure out if the tools you tried were successful.

Climate-related impacts are projected to occur across the US with increasing severity. Much funding and attention has been directed towards vulnerabiltiy assessments and adaptation. However, it is unclear how effective the assessments are at assessing vulnerability, and many do not consider social vulnerability. Research has shown that people who do not percieve themselves at risk, are less likely to adapt and are actually at a hieghtened risk. Currently, many people across the US still do not perceive themselves to be at risk of climate change impacts. People who do not perceive personal risk will be unmotivated to prepare for or adapt to such risks. Here, we map alignment between risk perception and future risk of climate change related impacts across the US. We find that certain counties where climate change risks are high, perception of risk remains low. In these regions of high mismatch between projected risk and perception of risk, we further investigate other factors of vulnerability which can exacerbate or alleviate actual risks, or perceptions of risk. This study is timely because it considers climate-related impacts such as drought, wildfire and heatwaves, that have and will continue to impact lives across the country. The findings of this study can inform managers of particularly vulnerable regions in the country where climate related risks are highly likely, but where risk is not perceived. It is in these regions that preparation is most critical and potentially most challenging. With limited resources available to improve adaptation and increase preparedness, relying only on risk projections may be an ineffective way to manage for climate impacts. Most vulnerability assessments focus on ecological impacts, and social vulnerability assessments rarely, if ever, consider risk perception as it pertains to sensitivity or adaptive capacity. These gaps highlight a critical need to identify where risk perception and mismatch of risk (mis)align.


## PSUEDOCODE!!
Before you get started programming all of the potential things you could do during the course of this demo. Lay out the steps you'll need to get there (see below for what I mean). You don't have to actually code these things (yet) just help us see how you're approaching the problem.
```{r}
#load the libraries - tell us which packages you're using and why

library(leaflet)
library(tidyr)
library(tigris) 
library(rgdal)
library(RColorBrewer)
library(sf)
library(raster)
library(sp)
library(fasterize)
library(mapview)
library(dplyr)
library(tmaptools)
library(ggplot2)
library(velox)


#load your data - which datasets do you need?

##Shapefile of Western counties
#Risk perception survey data
#Climate change risk data

#Organize the data - what form should the data be in? A list? how many elements, a data frame? how many rows and columns?

#Shapefile - 

#Analysis/processing step 1 - what are you hoping to do here, why? 

#Analysis/processing step 2 - what are you hoping to do here, why? 

#Check the outcomes? How will you know if your steps worked?

# 
```

## Introduce the packages
Given your psuedo code, where is the critical step? What packages and functions are you considering to help you complete this step? Why did you choose them? 

## Evaluate your choices
Use profiling and benchmarking to evaluate which of your options is likely to be the fastest. How does the syntax and/or ease of use of that function impact your decision of whether or not to use it? (For example, velox is much faster than raster, but it's less well documented and the syntax is strange to get used to).

## Show us your final product
Did you make a map? Let's see it. Did you plot some data that you extracted with raster? show us that plot. Did you have an idea of how the data should look after you were done processing it? Were you successful? What went wrong

## Reflect
Write a few sentences on what you learned from this exercise. How has your skill improved? What do you wish you understood better? What do you imagine your next steps to be?

Once you're done push the "knit" button to create the html page from your Rmarkdown document. If you've got questions, let me know!!