---
title: "Telemetry: Home Range Analysis"
author: "Brian Busby"
date: "4/21/2020"
output: html_document
---



## The Objective:  
 
  One objective of my research is to understand how Barn Owl behavioral patterns change in the presence of roads. While this will ultimately be a daunting and complex analysis, a good first step is to understand how Barn Owls are using the space around them and to what degree. One of the simpler estimates for habitat use is a home range analysis, which attempts to quantify the amount and location of land used by an individual. With this project I would like to successfully map out the home range of a Barn Owl that I tracked this winter using a GPS datalogger, and to do so efficiently. A final end product will hopefully be two maps, one of the owl's home range using a minimum convex polygon approach, and one using a kernel density estimator. 
  
## Pseudocode:

```{r}

#Load the required packages

#Import the Barn OWl GPS tracks as a .csv file

#Import the National Land Cover Dataset for Canyon County as a raster

#Align the projections for my tracks and the NLCD so the two can be plotted on top of eachother correctly

#Create a home range estimate using the minimum convex polygon approach and calculate area

#Determine the speed of the above process

#Create a home range estimate using a kernel density estimator and calculate area

#Determine the speed of the above process

#Repeat the above steps but with an alternate package

#Determine which package is more efficient 

#Create a nice visual displaying my results

```

## Packages Used:  

```{r, message=FALSE, warning=FALSE}

library(adehabitatHR)
# A telemetry package with a focus on home range estimations
library(amt)
# Our alternative home range estimation package. amt does not specialize in HRE as adehabitatHR does, however, it has a wide variety of uses for telemetry data. Therefore home ranges are not easily plottable from amt but we can compare area calculations
library(sp)
# to help me work with spatial objects
library(raster)
# to import and read our NLCD raster, which will serve as our background.
library(microbenchmark)
#to compare the speeds 
library(dplyr)
#for the tibble function needed to make tracks for the amt package
library(googledrive)
#for loading in the data

```

## Actual Code:


### 1. Data Preparation

```{r}
#first import your track data. In this case we only have one individual animal, so I can just read in the lat and long coordinates for each point. Since I'm just creating home ranges there is no need to read in time. 
#NOTE: a large step not pictured here is pre-processing your data as many telemetry units don't automatically compute this. See Joo et al. (2019) for more information.

trax <- read.csv("trackingdata.csv")
nlcd <- raster("NLCD1.tif")

head(trax)

#next we must convert the dataframe into spatial points so it can be read by the home range functions
trax2 <- SpatialPoints(trax)

#we also must assign it a coordinate reference system. This isn't necessary for the home range functions but it will make the area calculations simpler and will allow us to plot it with other background maps
#NOTE:for this to work the data must be organized first by longitude, then by latitute!
crs(trax2) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 

#read in an NLCD file fit to the shape of my study area. 
nlcd<-raster("NLCD1.tif")

#determine the projection
proj4string(nlcd)

#transform my tracks to match the projection of the NLCD
trax3<-spTransform(trax2, CRS(projargs = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0" ))

#visualize it
plot(nlcd, ext = extent(trax3)) +
plot(trax3, add = T)

```


### 2. MCP in adehabitatHR

```{r}

#create the home range estimation using minimal convex polygons. By electing 95 we have chosen to exclude our outermost 5% of points 
hr1 <- mcp(trax3, percent = 95)

#plot the two on top of eachother
plot(trax3, pch = 1) +
  plot(hr1, add = T)

#determine the area of this owls home range
as.data.frame(hr1)

#this value is in hectares, so 7.78 km2, sounds about right based on what I've read in the literature

sizes <- mcp.area(trax3, percent=seq(50, 100, by = 1))
#and a plot of how home range size would be effected based on what level I selected. Recall, I went with the pretty standard 95%

```


### 3. KUD in adehabitatHR

```{r}
#first create the kernel utilization density by inputting the points as well as the smoothing method which here is just the default and is based on my data
khr1 <- kernelUD(trax3, h="href")

#visualize it
image(khr1)

#now we can make another home range plot from this data
homerange <- getverticeshr(khr1)

plot(trax3, pch = 1) +
  plot(homerange, add = T)

```


### 4. MCP with amt

```{r}

#the first step is to begin formatting our data into a "track" which is what amt functions can read
df1 <- tibble(x = trax$long, y = trax$lat)

# Now we can create a track
tr1 <- make_track(df1, x, y, crs = sp::CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

#we'll transform the track to our NLCD projection to match our other data
transform_coords(tr1, sp::CRS("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0" ))

#caculate home range
hr2 <- hr_mcp(tr1)

#pull out the area from that calculation in m2
hr_area(hr2)

#convert it to km2
hr_area(hr2)/1000000
#nice. 7.6 is very close to the value of 7.8 we got using adehabitatHR

```

## Evaluation:  

```{r}
#first we see how quickly ade.time.mb can calculate home range in MCP style
ade.time.mb <- microbenchmark(mcp(trax3, percent = 95),
                                 times=100)

ade.time.mb
#well it changes slightly every time I run it but for my computer it hovers around a mean of 26 ms and a median of 21 ms

#now how fast does amt do the same process?
amt.time.mb <- microbenchmark(hr_mcp(tr1), times = 100)

amt.time.mb
#this seems to be a tad bit faster with a mean of 14 ms and median of 10
```

  From a quick microbenchmarking assessment it appears that the amt package is about twice as fast as adehabitatHR at calculating area of home range using minimum convex polygons. That being said they both produce results extremely quickly given that thousands of coordinates are inputted to create these calculations. 

  If all you needed was home range area calculations and you had hundreds of different animals with tens of thousands of coordinates for each then you might actually notice a difference in speed by using amt. However, if you want to further explore home ranges beyond area calculations and you don't have an absolutely gargantuan data set then adehabitatHR is probably the best bet. The package is designed for home range work, the functions are much more intuitive than amt, and it's very easy to plot and create visuals of home ranges. It doesn't perform quite as fast as amt but if speed isn't an issue its definitely the package to go with.


## Final Product:

```{r}

#lets create some nice plots with our findings from adehabitatHR
#first, lets compare the KUD approach to the MCP approach by overlaying the two plots

#first we need some custom transparent colors so the overlay works
mycol <- rgb(0, 0, 255, max = 255, alpha = 50)
mycol1 <- rgb(0, 255, 0, max = 255, alpha = 50)

#and plot it
plot(trax3, pch = 1) +
  plot(hr1, add = T, col = mycol1) +
plot(homerange, col = mycol, add = T)
#green = MCP, blue = KUD

#Now lets make a nice graphic of both our MCP and KUD home ranges that include the data points and a raster NLCD background

#MCP:
plot(nlcd, ext = extent(trax3)) +
plot(trax3, add = T) +
  plot(hr1, col = mycol1, add = T)

#KUD:
plot(nlcd, ext = extent(trax3)) +
plot(trax3, add = T) +
  plot(homerange, col = mycol, add = T)

```


## Reflection:

  The big thing I learned was that home ranges are not as difficult to calculate as I anticipated once you have the proper tools in R. This project has helped me get better at managing coordinates in R, assigning and changing projections, doing home range calculations, and using plot to create visuals of my work. I still wish I was a little better at manipulating plot to get the right extent to my maps and that I could incorporate ggmap to create backgrounds better than just the NLCD. My next steps will likely be adding the third dimension, time, to my coordinates and starting to use these packages, particularly amt, to breakdown movement behavior in my Barn Owls. 
  

### Literature Cited:

Joo, R., M.E. Boone, T.A. Clay, S.C. Patrick, S. Clusella-Trullas, and M. Basille. 2019. Navigating through the R packages for movement. Journal of Animal Ecology 89:248 - 267
