---
title: "Interactive COVID-19 Chart for Idaho"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(dygraphs)
library(timetk)
library(here)
library(leaflet)
library(highcharter)
require(data.table)
require(tidyr)
require(ggplot2)
require(lubridate)
require(rgdal)
require(raster)
require(broom)
require(maptools)
require(rgeos)

```

Dashboard
=====================================


```{r load data, message=FALSE, warning=FALSE, include=FALSE}
#### Load datasets ####
#these datasets are on github so you can just put the link in to call up the file location on github
#This is the data forked from the John Hopkins repo at https://github.com/CSSEGISandData/COVID-19.
FullTimeSer_conf <- fread("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv", header = T)

#Preparing the data read in above by converting to long form and truncating dates
IdTimeSer_conf <- FullTimeSer_conf[Province_State == "Idaho",]
IdLong_conf <- as.data.table(gather(data = IdTimeSer_conf, key = Date, value = Cases, 12:ncol(IdTimeSer_conf)))
IdLong_conf[, Date := as.Date(mdy(Date, tz = "America/Boise"))]
IdLong_conf <- IdLong_conf[Date == max(Date) & !Admin2 == "Out of ID" & !Admin2 == "Unassigned"]

#Creating cumulative Idaho cases data table
IdSum_conf <- as.data.table(aggregate(IdLong_conf$Cases~              IdLong_conf$Admin2+IdLong_conf$Lat+IdLong_conf$Long_, FUN = sum))
setnames(IdSum_conf, c("name","lat","lon", "z"))
```


Column {data-width=300}
-----------------------------------------------------------------------

**Introduction**

This dashboard is an example for how to show time-series data in an interactive form. Specially, this project is focused on COVID-19 cases in the state of Idaho. A flexdashboard is a helpful tool for displaying this information because it can be used to make data visually appealing and more digestable. 

The goal of this flexdashboard is to compare two different mapping packages and assess the pros/cons to these packages. 

**Packages for this project**
Given your psuedo code, where is the critical step? What packages and functions are you considering to help you complete this step? Why did you choose them? 

Column {data-width=650}
-----------------------------------------------------------------------

### Chart A

```{r}
# we are going to use the highcharter package to display information

# let's pull up a map of idaho, at the county level
#hcmap("countries/us/us-id-all") %>%
  #hc_title(text = "Idaho")

# lets see what our data looks like

glimpse(IdSum_conf)
# need to convert cases to double
IdSum_conf$z <- as.double(IdSum_conf$z)

# I use glimpse to make sure my data is in the exact format the package needs: columns are "name", "lat", "long
glimpse(IdSum_conf)


# lets plot our data on the map
hcmap("countries/us/us-id-all", showInLegend=FALSE) %>% 
  hc_add_series(data = IdSum_conf, type = "mapbubble",
                name = "Confirmed COVID-19 Cases", maxSize = '10%') %>% 
  hc_mapNavigation(enabled=TRUE)
  
```

### Chart B

```{r leaflet}

# create leaflet map

id_confirm_map <- leaflet(IdSum_conf) %>% #this calls up the package
                  addProviderTiles("CartoDB.Positron") %>% # this load the basemap type
                  addCircleMarkers(lng=~lon, # this is the specifications for the marker
                                   lat=~lat,
                                   popup=~head(paste0(IdSum_conf$name, # this tells is what information to display and how!
                                                      "<br/>Confirmed Cases:", 
                                                      IdSum_conf$z)))
id_confirm_map
```

Background
=====================================

This dashboard was created to help explore two different ways of displaying COVID-19 data. 

# This is my workflow
```{r}
#load the libraries - tell us which packages you're using and why
# tidyverse- useful package for data management
# flexdashboard- necessary for making a flexdashboard
# dygraphs- makes cool charts
# here- specifies how to use data
# leaflet- mapping mapckage
# highcharter- mapping package

#load your data - which datasets do you need?
# I am getting my data from 2019 Novel Coronavirus COVID-19 (2019-nCoV) Data Repository by Johns Hopkins Center for Systems Science and Engineering 
# I am using two datasets: one for confirmed and one for deaths
#Organize the data - what form should the data be in? A list? how many elements, a data frame? how many rows and columns?
# The data comes in a time series format-- I need to take the data and create a summation of all cases per county. 
#Analysis/processing step 1 - what are you hoping to do here, why? 
# Once I have the summary's I am going to plot them on a map using highcharter

#Analysis/processing step 2 - what are you hoping to do here, why? 
# Then I will plotting them on map using leaflet

#Check the outcomes? How will you know if your steps worked?
# I will know if I have two functioning maps of Idaho with accurate totals!
```

## Evaluate your choices
Use profiling and benchmarking to evaluate which of your options is likely to be the fastest. How does the syntax and/or ease of use of that function impact your decision of whether or not to use it? (For example, velox is much faster than raster, but it's less well documented and the syntax is strange to get used to).

## Show us your final product
Did you make a map? Let's see it. Did you plot some data that you extracted with raster? show us that plot. Did you have an idea of how the data should look after you were done processing it? Were you successful? What went wrong

## Reflect
Write a few sentences on what you learned from this exercise. How has your skill improved? What do you wish you understood better? What do you imagine your next steps to be?

Once you're done push the "knit" button to create the html page from your Rmarkdown document. If you've got questions, let me know!!

The data used to build the analysis is taken from the data repository for the 2019 Novel Coronavirus Visual Dashboard, operated by the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE). Their GitHub repo is linked [here](https://github.com/CSSEGISandData/COVID-19).

This dashboard was created by Tara Pozzi. 
